---
title: "SDoumbia_Assignment10_MarkovChains/RandomWalks"
author: "Souleymane Doumbia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if
(a) he bets 1 dollar each time (timid strategy).
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
(c) Which strategy gives Smith the better chance of getting out of jail?

Given:

- Smith starts with 1 dollar.

- He needs 8 dollars to get out of jail.

- Each bet can result in winning or losing the bet amount.

- The probability of winning a bet is 0.4, and the probability of losing is 0.6.


## Part (a): Timid Strategy

Under the timid strategy, the recursive formula for the probability \(P(n)\) when betting 1 dollar each time is:

\[P(n) = 0.4P(n+1) + 0.6P(n-1)\]

with \(1 \leq n \leq 7\), and the boundary conditions \(P(8) = 1\) and \(P(0) = 0\).

```{r Timid_Strategy}
p_win <- 0.4
n_states <- 9

# Initializing probability vector with zeros
P <- numeric(n_states)

P[9] <- 1 

# Calculating probabilities for the timid strategy
for (i in (n_states-1):1) {
  if (i == 1) {  
    P[i] <- 0
  } else {
    P[i] <- p_win * P[i + 1] + (1 - p_win) * P[i - 1]
  }
}

# Probability of reaching $8 starting from $1
P_timid <- P[2]
P_timid

```
After calculating, we find that the probability of reaching $8 starting from $1 under the timid strategy is approximately 0.164%.

## Part (b): Bold Strategy

For the bold strategy, Smith bets as much money as he can without exceeding the amount he needs to reach 8 dollars. The amount bet is \(min(n, 8-n)\) where \(n\) is the current amount of money Smith has. The recursive formula adjusts for these variable bet sizes:

\[P(n) = 0.4P(n + min(n, 8-n)) + 0.6P(n - min(n, 8-n))\]

With the same boundary conditions as the timid strategy (\(P(8) = 1\) and \(P(0) = 0\)), we solve for \(P(n)\) for each state from \(n = 7\) down to \(n = 1\) considering the bold strategy of betting.

```{r Bold_Strategy}
p_win <- 0.4
n_states <- 9

P_bold <- numeric(n_states)

P_bold[9] <- 1 # Winning condition

# Iterating through each state to calculate probabilities
for (i in 8:1) {
  if (i < 5) {
    # If Smith has less than $5, bet all his money
    # Since he can't lose more than he bets, we only consider the winning scenario.
    P_bold[i] <- p_win * P_bold[i*2]
  } else {
    # If Smith has $5 or more, bet only what's needed to reach $8
    P_bold[i] <- p_win * P_bold[9] + (1 - p_win) * P_bold[i - (8 - i)]
  }
}

# Probability of reaching $8 starting from $1 under the bold strategy
P_bold_1 <- P_bold[2]
P_bold_1

```
Applying this strategy, the probability of Smith winning 8 dollars before going broke is found to be approximately 6.4%.


## Part (c): Comparison

- **Timid Strategy Probability:** 0.164%
- **Bold Strategy Probability:** 6.4%

Thus, the **bold strategy** significantly increases Smith's chances of achieving his goal of accumulating 8 dollars and securing his release from jail.



* * *


