---
title: "SDoumbia_Assignment13_UnivariateMultivariateCalculus"
author: "Souleymane Doumbia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1: 
Using integration by substitution to solve the integral:
$$
\int 4e^{-7x} \, dx
$$

Let $u = -7x$, therefore $du = -7dx$ or $dx = -\frac{1}{7}du$. Substituting into the integral, we get:
$$
\int 4e^{u} \cdot \frac{du}{-7} = -\frac{4}{7} \int e^{u} \, du = -\frac{4}{7}e^{u} + C
$$
Now, since $u = -7x$, we replace $u$ back with $-7x$:
   $$
   -\frac{4}{7} e^{u} + C = -\frac{4}{7} e^{-7x} + C
   $$

Thus, the integral $\int 4e^{-7x} \, dx$ evaluates to:
$$
\boxed{-\frac{4}{7} e^{-7x} + C}
$$

## Problem 2: 
To solve for the function \(N(t)\) that estimates the level of contamination in a pond, we start by setting up the differential equation based on the given rate of change of contamination. The rate of change \(dN/dt\) is described as follows:


\[
\frac{dN}{dt} = -\frac{3150}{t^4} - 220
\]

where \( t \) is the number of days since the treatment began. 

### Let s find \(N(t)\):

The function \(N(t)\) can be found by integrating the rate of change:

\[
N(t) = \int \left(-\frac{3150}{t^4} - 220\right) \, dt
\]

This integral can be computed as follows:

\[
N(t) = \int -\frac{3150}{t^4} \, dt - \int 220 \, dt
\]

Integrating term-by-term, we get:

\[
N(t) = 1050 \frac{1}{t^3} - 220t + C
\]

where \( C \) is the integration constant.

### Applying the Initial Condition

It is given that after 1 day the level of contamination was 6530 bacteria per cubic centimeter. This initial condition helps us determine \(C\):

\[
1050 \cdot \frac{1}{1^3} - 220 \cdot 1 + C = 6530
\]

\[
1050 - 220 + C = 6530
\]

\[
C = 6530 - 1050 + 220 = 5700
\]

Thus, the function that models the level of contamination \( N(t) \) is:

\[
N(t) = 1050 \frac{1}{t^3} - 220t + 5700
\]


## Problem 3
To find the total area of the red rectangles under the line described by the function \( f(x) = 2x - 9 \) from \( x = 4.5 \) to \( x = 8.5 \), we calculate the definite integral of the function over this range. The integral will give us the area between the line and the x-axis.

**Function Definition**:
   The line equation is given by:
   $$
   f(x) = 2x - 9
   $$
   We need to integrate this function from \( x = 4.5 \) to \( x = 8.5 \).

**Calculating the Integral**:
   The integral of \( f(x) \) from \( x = 4.5 \) to \( x = 8.5 \) is:
   $$
   \int_{4.5}^{8.5} (2x - 9) \, dx
   $$
   Solving this integral:
   $$
   \int (2x - 9) \, dx = x^2 - 9x + C
   $$
   Applying the limits of integration:
   $$
   \left[ x^2 - 9x \right]_{4.5}^{8.5} = (8.5^2 - 9 \times 8.5) - (4.5^2 - 9 \times 4.5)
   $$
   $$
   = (72.25 - 76.5) - (20.25 - 40.5)
   $$
   $$
   = -4.25 + 20.25 = 16
   $$


## Problem 4

To find the area of the region bounded by the curves given by the equations \(y = x^2 - 2x - 2\) and \(y = x + 2\), we will follow these steps:

**Setting the Equations Equal to Find Intersection Points**:
   $$
   x^2 - 2x - 2 = x + 2
   $$
   solving for \(x\):
   $$
   x^2 - 3x - 4 = 0
   $$
   Factorizing the quadratic equation:
   $$
   (x - 4)(x + 1) = 0
   $$
   This gives the solutions \(x = 4\) and \(x = -1\).

**Determining which Function is Above the Other**:
   Substituting values between \(x = -1\) and \(x = 4\) into both functions to determine which is above the other. We observed that \(y = x + 2\) lies above \(y = x^2 - 2x - 2\) in this interval.

**Setting Up the Integral for the Area**:
  
   Here, \(f(x) = x + 2\) and \(g(x) = x^2 - 2x - 2\). 
   $$
   A = \int_{-1}^{4} [(x + 2) - (x^2 - 2x - 2)] \, dx = \int_{-1}^{4} [-x^2 + 3x + 4] \, dx
   $$

**Computing the Integral**:
   $$
   A = \left[ -\frac{1}{3}x^3 + \frac{3}{2}x^2 + 4x \right]_{-1}^{4}
   $$
   $$
   A = \left( -\frac{1}{3}(4)^3 + \frac{3}{2}(4)^2 + 4(4) \right) - \left( -\frac{1}{3}(-1)^3 + \frac{3}{2}(-1)^2 + 4(-1) \right)
   $$
   Simplifying to get:
   $$
   A = \left( -\frac{64}{3} + 24 + 16 \right) - \left( \frac{1}{3} + \frac{3}{2} - 4 \right)
   $$
   $$
   A = -\frac{64}{3} + 40 + \frac{1}{3} - \frac{3}{2} + 4
   $$
   \[
   A = -\frac{64}{3} + \frac{1}{3} - \frac{4.5}{3} = -\frac{64 + 4.5 - 1}{3} + 44= -\frac{67.5}{3} = 40 -22.5
   \]

Thus,  
\[
Area = 21.5
\]


## Problem 5

### Economic Order Quantity (EOQ) Model
To determine the optimal lot size and number of orders per year that minimize inventory costs for a beauty supply store, we use the Economic Order Quantity (EOQ) model. 

The EOQ model is defined by the following parameters and formula:

- **Annual Demand (\(D\))**: The total units demanded over the course of a year.

- **Holding Cost (\(H\))**: The cost to hold one unit in inventory for a year.

- **Ordering Cost (\(S\))**: The fixed cost associated with placing an order.

The formula for EOQ is given by:
$$
EOQ = \sqrt{\frac{2DS}{H}}
$$
Where:
- \(D\) is the annual demand,
- \(S\) is the ordering cost per order,
- \(H\) is the holding cost per unit per year.

### Given Values

- Annual demand \(D = 110\) flat irons.

- Ordering cost \(S = \$8.25\) per order.

- Holding cost \(H = \$3.75\) per unit per year.

```{r eoq_calculation}
# Define the parameters
D <- 110  # Annual demand
S <- 8.25  # Ordering cost per order
H <- 3.75  # Holding cost per unit per year

# Calculate EOQ
EOQ <- sqrt((2 * D * S) / H)
EOQ
```

### Calculate Number of Orders Per Year
```{r Number_of_Orders_Per_Year}
number_of_orders <- D / EOQ
number_of_orders
```


### Calculate Total Inventory Cost
The total inventory cost (TC) when ordering at the EOQ can be calculated as:

\[
TC = \left(\frac{D}{Q} \times S\right) + \left(\frac{Q}{2} \times H\right)
\]

where \(Q\) is the order quantity, which is the EOQ in this case.

```{r Total_Inventory_Cost}
total_cost <- (D / EOQ) * S + (EOQ / 2) * H
total_cost
```


## Problem 6
Use integration by parts to solve the integral below:
\[ \int x \ln(9x) \, dx \]


### Integration by Parts

integration by parts to solve:

\[ \int \ln(9x) \cdot x \, dx \]

We choose:
\[ u = \ln(9x), \quad dv = x \, dx \]

We integrate to find \( du \) and \( v \):
\[ du = \frac{1}{x} \, dx, \quad v = \frac{x^2}{2} \]

Using the integration by parts formula:
\[ \int u \, dv = uv - \int v \, du \]

We substitute \( u \), \( dv \), \( du \), and \( v \):
\[ \int \ln(9x) \cdot x \, dx = \frac{x^2}{2} \ln(9x) - \int \frac{x^2}{2} \cdot \frac{1}{x} \, dx \]
\[ = \frac{x^2}{2} \ln(9x) - \int \frac{x}{2} \, dx \]
\[ = \frac{x^2}{2} \ln(9x) - \frac{x^2}{4} + C \]

Where \( C \) is a constant. 


## Problem 7
Determine whether the function \( f(x) \) is a probability density function on the interval \([1, e^6]\). If not, determine the value of the definite integral. The function is given by
\[ f(x) = \frac{1}{6x}. \]

#### Checking if a Function is a Probability Density Function

To determine whether the function \( f(x) = \frac{1}{6}x \) is a probability density function (PDF) on the interval \([1, e^6]\), we need to verify two conditions:

1. The function must be non-negative over the interval.
2. The integral of the function over the interval must equal 1.

#### Step 1: Check Non-Negativity

For \( f(x) = \frac{1}{6}x \), since \( x \geq 1 \) over the interval \([1, e^6]\), \( f(x) \) is non-negative.

#### Step 2: Check if the Integral Equals 1

We calculate the integral of \( f(x) \) from 1 to \( e^6 \):

\[
\int_{1}^{e^6} \frac{1}{6}x \, dx = \frac{1}{6} \int_{1}^{e^6} x \, dx
\]

This integrates to:

\[
= \frac{1}{6} \left[\frac{x^2}{2}\right]_{1}^{e^6} = \frac{1}{6} \left(\frac{(e^6)^2}{2} - \frac{1^2}{2}\right) = \frac{1}{6} \left(\frac{e^{12}}{2} - \frac{1}{2}\right)
\]

\[
= \frac{e^{12} - 1}{12}
\]

Since this value is not equal to 1, \( f(x) \) is not a PDF.

#### Correcting \( f(x) \) to Be a PDF

To correct \( f(x) \), we need to find a constant \( c \) such that:

\[
c \int_{1}^{e^6} x \, dx = 1
\]

Solving for \( c \):

\[
c \cdot \frac{e^{12} - 1}{2} = 1 \Rightarrow c = \frac{2}{e^{12} - 1}
\]

Thus, the corrected function is:

\[
\boxed{f(x) = \frac{2}{e^{12} - 1}x}
\]

This function \( f(x) \) now qualifies as a PDF over the interval \([1, e^6]\).



* * *

