---
title: 'Assignment 2: Trace / Determinant / Factorization'
author: "Souleymane Doumbia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Matrix)
```

###  1.  Problem set 1

#### (1) Showing $A^TA \neq AA^T$ in General

Given $A$ as an $m \times n$ matrix, $A^T$ is an $n \times m$ matrix. 

The product $A^TA$ results in an $n \times n$ matrix, while $AA^T$ results in an $m \times m$ matrix.

##### Proof Using an Example:

Let's take a matrix $A$:

$$
A = \begin{pmatrix}
1 & 2 \\
3 & 4 \\
\end{pmatrix}
$$

The transpose, $A^T$, is:

$$
A^T = \begin{pmatrix}
1 & 3 \\
2 & 4 \\
\end{pmatrix}
$$

Calculating $A^TA$:

$$
A^TA = \begin{pmatrix}
1 & 3 \\
2 & 4 \\
\end{pmatrix}
\begin{pmatrix}
1 & 2 \\
3 & 4 \\
\end{pmatrix}
= \begin{pmatrix}
10 & 14 \\
14 & 20 \\
\end{pmatrix}
$$

Calculating $AA^T$:

$$
AA^T = \begin{pmatrix}
1 & 2 \\
3 & 4 \\
\end{pmatrix}
\begin{pmatrix}
1 & 3 \\
2 & 4 \\
\end{pmatrix}
= \begin{pmatrix}
5 & 11 \\
11 & 25 \\
\end{pmatrix}
$$

Thus, $A^TA \neq AA^T$, demonstrating that in general, $A^TA \neq AA^T$.


#### (2) Conditions for $A^TA = AA^T$

For the equality $A^TA = AA^T$ to hold, the matrix $A$ must satisfy certain conditions. This is true for:

- **Symmetric Matrices**: Where $A = A^T$. For such matrices, it's common that $A^TA = AA^T$ due to their symmetry.
- **Orthogonal Matrices**: Which satisfy $A^TA = I$, where $I$ is the identity matrix. Since $A^T = A^{-1}$ for orthogonal matrices, both $A^TA$ and $AA^T$ yield the identity matrix, indicating that $A^TA = AA^T$ in this case.





###  2.  Problem set 2

**Here is R function for LU decomposition of a square matrix**
```{r luDecomposition_function_square_matrix}

luDecomposition <- function(A) {
  if (!is.matrix(A)) {
    stop("Input must be a matrix.")
  }
  
  n <- nrow(A)
  if (ncol(A) != n) {
    stop("Matrix must be square.")
  }
  
  L <- diag(1, n) # Initializing L as an identity matrix of size n
  U <- A # Copying A into U for transformation
  
  for (j in 1:(n-1)) {
    for (i in (j+1):n) {
      if (U[j, j] == 0) {
        stop("Zero pivot encountered.")
      }
      L[i, j] <- U[i, j] / U[j, j]
      U[i, ] <- U[i, ] - L[i, j] * U[j, ]
    }
  }
  
  return(list(L = L, U = U))
}

```

**An example of Application using the LU function above:**
```{r Application_of_luDecomposition_function}

A <- matrix(c(1, 2, 1, 3,
              3, 8, 1, 5,
              0, 4, 1, 9,
              4, 11, 3, 0), byrow = TRUE, nrow = 4)

result <- luDecomposition(A)

print("L Matrix:")
print(result$L)

print("U Matrix:")
print(result$U)
```

* * *
