---
title: "Assignment 3"
author: "Souleymane Doumbia"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem set 1

**1. What is the rank of the matrix A?**

\[ A = \begin{pmatrix}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3
\end{pmatrix} \]

Matrix A is a 4x4 matrix, then the rank is 4.


**2. Maximum and Minimum Rank of a Matrix**

For a matrix of size \(m \times n\) with \(m > n\), the maximum possible rank is \(n\), which corresponds to the maximum number of linearly independent column vectors in the matrix. The minimum rank, assuming the matrix is non-zero, is 1, because even a single non-zero element in the matrix indicates a non-zero determinant for some \(1 \times 1\) submatrix.

**3. What is the rank of Matrix B**


\[ B = \begin{pmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2
\end{pmatrix} \]

Matrix B is a 3x3 matrix, then the rank is 3.



### Problem set 2

\[ A = \begin{pmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6
\end{pmatrix} \]


**Computing Eigenvalues and Eigenvectors:**

```{r Computing Eigenvalues and Eigenvectors}
# Defining the matrix A
A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow = 3, byrow = TRUE)

# Calculating eigenvalues and eigenvectors
eigen_A <- eigen(A)

eigen_A$values
eigen_A$vectors
```



**Characteristic Polynomial of Matrix A**

---
title: "Characteristic Polynomial"
output: html_document
---

## Characteristic Polynomial of Matrix A


The eigenvalues are 1, 4, and 6. The characteristic polynomial \( p(\lambda) \) is given by the determinant of \( A - \lambda I \):

\[p(\lambda) = \det(A - \lambda I) = (\lambda - 1)(\lambda - 4)(\lambda - 6)\]


### Calculating the Characteristic Polynomial in R

```{r}
# Assuming matrix A is already defined as before
# Calculate the characteristic polynomial coefficients
char_poly <- function(lambda) {
  det(matrix(c(1 - lambda, 2, 3, 
               0, 4 - lambda, 5, 
               0, 0, 6 - lambda), nrow = 3, byrow = TRUE))
}

# Display the characteristic polynomial for each eigenvalue
eigenvalues <- c(1, 4, 6)
sapply(eigenvalues, char_poly)
```

My problem here so far, how to ask R to express the result as an expression of \( \lambda\).







