---
title: "SDoumbia Chap. 8.8, Exercise 9, P. 496"
author: "Souleymane Doumbia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Finding the $n^{th}$ Taylor Series for $f(x) = e^{-x}$ at $c = 0$

## Formula Derivation

To find the Taylor series for the function $f(x) = e^{-x}$ centered at $c = 0$, we first compute the derivatives of $f(x)$ at $x = 0$. The Taylor series expansion of a function $f(x)$ around a point $c$ is given by:

\[
f(x) = \sum_{n=0}^{\infty} \frac{f^{(n)}(c)}{n!}(x-c)^n
\]

For $f(x) = e^{-x}$ and $c = 0$, we observe the following pattern in the derivatives:

- $f(x) = e^{-x}$
- $f(0) = 1$
- $f'(x) = -e^{-x}$
- $f'(0) = -1$
- $f''(x) = e^{-x}$
- $f''(0) = 1$
- Continuing this pattern, the $n$-th derivative evaluated at $x = 0$ is:

\[
f^{(n)}(0) = (-1)^n
\]

Hence, the formula for the Taylor series becomes:

\[
f(x) = \sum_{n=0}^{\infty} \frac{(-1)^n}{n!}x^n
\]

## Implementation Using Pracma package

We will now use R to compute the Taylor series expansion of $f(x) = e^{-x}$ at $c = 0$ up to the 6th order, using the `pracma` package.

```{r}
# Load the pracma package; 
library(pracma)

# Define the function
myf <- function(x) exp(-x)

# Compute the Taylor series
taylor_series <- taylor(myf, x0 = 0, n = 6)

# Print the result
print(taylor_series)
```

