---
title: "Hw15 Functions Of Several Variables"
author: "Souleymane Doumbia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# $\underline{PROBLEM  1}$

Given points: (5.6, 8.8), (6.3, 12.4), (7, 14.8), (7.7, 18.2), (8.4, 20.8)

We use the method of least squares to find the best-fitting line for the given data points. 

### Linear Regression Equation
The linear regression equation is given by:

\[ y = mx + b \]

where:
- \( y \) is the dependent variable,
- \( x \) is the independent variable,
- \( m \) is the slope of the line,
- \( b \) is the y-intercept.

### Calculation of Slope and Intercept
The slope \( m \) and intercept \( b \) are calculated using:

\[ m = \frac{N(\sum xy) - (\sum x)(\sum y)}{N(\sum x^2) - (\sum x)^2} \]
\[ b = \frac{(\sum y) - m(\sum x)}{N} \]

where:
- \( N \) is the number of data points,
- \( \sum xy \) is the sum of the product of each pair of \( x \) and \( y \),
- \( \sum x \), \( \sum y \) are the sums of \( x \) and \( y \) values respectively,
- \( \sum x^2 \) is the sum of the squares of \( x \) values.

### R Code for Calculation

```{r regression_calculation}
# Defining the data points
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

# Calculating for slope (m) and intercept (b)
N <- length(x)
sum_x <- sum(x)
sum_y <- sum(y)
sum_xy <- sum(x*y)
sum_x2 <- sum(x^2)

# Slope
m <- (N * sum_xy - sum_x * sum_y) / (N * sum_x2 - sum_x^2)

# Intercept
b <- (sum_y - m * sum_x) / N

# Rounding to the nearest hundredth
m <- round(m, 2)
b <- round(b, 2)

# Output the slope and intercept
c(m = m, b = b)
```


# $\underline{PROBLEM  3}$

## Revenue Analysis for Different Product Pricings

## Defineing the Revenue Function

We have:

- **House Brand**: Estimated units sold \( = 81 - 21x + 17y \)
- **Name Brand**: Estimated units sold \( = 40 + 11x - 23y \)

Given these estimates, the revenue function, representing the total sales from each product, is:

\[ R(x, y) = x(81 - 21x + 17y) + y(40 + 11x - 23y) \]

Expanding and simplifying, we have:

\[ R(x, y) = -21x^2 + 28xy - 23y^2 + 81x + 40y \]

This function calculates the total revenue based on prices \( x \) and \( y \) for the house and name brands, respectively.

## Computing Revenue for Specific Prices

For prices:
- House brand \( x = \$2.30 \)
- Name brand \( y = \$4.10 \)

The revenue can be calculated by substituting these prices into the revenue function:

\[ R(2.30, 4.10) = -21(2.30)^2 + 28(2.30)(4.10) - 23(4.10)^2 + 81(2.30) + 40(4.10) \]


Then, this will yield the total revenue from selling the products at these prices, which is:
```{r calculate_revenue}
x <- 2.30
y <- 4.10
R <- -21*x^2 + 28*x*y - 23*y^2 + 81*x + 40*y
R_rounded <- round(R, 2)
R_rounded
```



# $\underline{PROBLEM  4}$

## Minimizing Production Costs

A company must produce 96 units weekly between two plants. The cost function is:

\[ C(x, y) = 16x^2 + 16y^2 + 7x + 25y + 700 \]

where \( x \) and \( y \) are the units produced in Los Angeles and Denver, respectively.


Our objective is to determine how many units should be produced at each location to minimize the total cost, given the constraint \( x + y = 96 \).


Using the method of Lagrange multipliers, we set up the Lagrangian as follows:

\[ L(x, y, \lambda) = 16x^2 + 16y^2 + 7x + 25y + 700 + \lambda (96 - x - y) \]

### Calculating the Derivatives

Derivatives of the Lagrangian are set to zero to solve for \( x \), \( y \), and \( \lambda \):

\[ \frac{\partial L}{\partial x} = 32x + 7 - \lambda = 0 \]
\[ \frac{\partial L}{\partial y} = 32y + 25 - \lambda = 0 \]
\[ \frac{\partial L}{\partial \lambda} = 96 - x - y = 0 \]

### Solving the System of Equations

```{r solve_system}
library(nleqslv)

# System of equations based on the Lagrangian derivatives
system_eq <- function(z) {
  x <- z[1]
  y <- z[2]
  lambda <- z[3]

  # System of equations
  eq1 <- 32*x + 7 - lambda
  eq2 <- 32*y + 25 - lambda
  eq3 <- 96 - x - y

  c(eq1, eq2, eq3)
}

# Initial guesses for x, y, and lambda
z_start <- c(x = 48, y = 48, lambda = 100)

# Solving the system
solution <- nleqslv(x = z_start, fn = system_eq)

# Extracting solutions
x_opt <- solution$x[1]
y_opt <- solution$x[2]

# Output the results
c("Units to produce in Los Angeles" = x_opt, "Units to produce in Denver" = y_opt)
```


# $\underline{PROBLEM  5}$

We are given the double integral over the region \( R \) defined by \( 2 \leq x \leq 4 \) and \( 2 \leq y \leq 4 \) for the function \( e^{8x+3y} \).

The function to integrate is \( e^{8x + 3y} \), and the region of integration is defined by the ranges of \( x \) and \( y \).

\[ R: 2 \leq x \leq 4, 2 \leq y \leq 4 \]

## Step-by-Step Calculation

### Step 1: 

We start by setting up the double integral:

\[ \int_{2}^{4} \int_{2}^{4} e^{8x+3y} \, dy \, dx \]

### Step 2: Integrate with Respect to \( y \)

First, integrate with respect to \( y \), treating \( x \) as a constant:

\[ \int_{2}^{4} e^{8x+3y} \, dy = e^{8x} \int_{2}^{4} e^{3y} \, dy \]

The integration with respect to \( y \) yields:

\[ = e^{8x} \left[ \frac{e^{3y}}{3} \right]_{y=2}^{y=4} \]
\[ = e^{8x} \left( \frac{e^{12}}{3} - \frac{e^6}{3} \right) \]
\[ = \frac{e^{8x}}{3} (e^{12} - e^6) \]

### Step 3: Integrate with Respect to \( x \)

Now, integrating the result with respect to \( x \):

\[ \int_{2}^{4} \frac{e^{8x}}{3} (e^{12} - e^6) \, dx \]
\[ = \frac{e^{12} - e^6}{3} \int_{2}^{4} e^{8x} \, dx \]

The integration with respect to \( x \) is:

\[ = \frac{e^{12} - e^6}{3} \left[ \frac{e^{8x}}{8} \right]_{x=2}^{x=4} \]
\[ = \frac{e^{12} - e^6}{3} \left( \frac{e^{32}}{8} - \frac{e^{16}}{8} \right) \]
\[ = \frac{(e^{12} - e^6) (e^{32} - e^{16})}{24} \]


Therefore, the total value of the double integral over the specified region \( R \) is:

\[ \frac{(e^{12} - e^6) (e^{32} - e^{16})}{24} \]


### Estimating the result:
```{r calculate_integral}
# Evaluate the final expression
e <- exp(1) # Base of the natural logarithm
result <- (e^12 - e^6) * (e^32 - e^16) / 24
result
```



* * *


